---
title: "NARX Model Guide"
author: "Farzad"
date: "15 August 2016"
output: rmarkdown::html_vignette
#  html_document:
#    self_contained: false
#    lib_dir: libs
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{NARX Model Guide}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("mltsp")
library("fpp")
library("e1071")
```

## Setup

For the rest of this guide, we will use time-series `ausbeer` from package `fpp`:

```{r includes}
library("fpp")

beer_train = head(ausbeer, -20)
beer_test = head(tail(ausbeer, 20), 10)
```

## NARX Model

A simple NARX model can be crafted using `narx`. 
A model with 

 * differencing order `d=0`
 * auto-regressive order `p=2`
 * seasonal frequency `freq=4`
 * seasonal differencing order `D=1`
 * seasonal order `P=1`
 * `svm` learner
 
can be instantiated as follows:

```{r buildnarx}
library("mltsp")
library("e1071")
spec = build_narx(svm, p=2, d=0, P=1, D=1, freq=frequency(ausbeer))
model = narx(spec, beer_train)

model
```

or alternatively:
```{r buildnarx2}
model = narx(beer_train, svm, p=2, d=0, P=1, D=1, freq=frequency(ausbeer))
```

## Duplicating the model

One could re-instantiate the model for use with another dataset using
```{r buildnarx3}
beer_train2 = head(ausbeer, -10)

model2 = narx(model, beer_train2)
```

## Prediction

Prediction uses either `forecast` (compatible with package `forecast`):
```{r testnarx}
fcst = forecast(model, h = 10)
plot(fcst)
lines(beer_test, col="red")
```

or use `predict` if you want to also supply new data but use the same model:
```{r testnarx2}
beer_train2 = head(ausbeer, -10) # this is using data from another future!
beer_test2 = tail(ausbeer, 10) # this is using data from another future!

fcst2 = predict(model, beer_train2, h = 10)

plot(ausbeer)
lines(beer_test2, col="red")
lines(fcst2, col="blue")
```

Alternatively, one could reuse the model using
```{r testnarx3}
model2 = narx(model, beer_train2)
fcst3 = forecast(model2, h = 10)

plot(fcst3)
lines(fcst2, col="red")
```


## Using exogenous data

Use `xreg` parameter as exogenous data. This example tries to forecast a random walk. Without xreg, this should be almost impossible:

```{r xreg}
set.seed(0)

tstamps = seq(as.Date("2000-01-01"), length.out = 110, by='day')
x = xts(runif(length(tstamps)), tstamps)
xreg = 1 - 0.5 * x
yreg = xts(runif(110), tstamps)

colnames(xreg) = colnames(yreg) = "xreg"

# training and testing data
x_train = head(x, 100)
x_test = tail(x, 10)
ind_test = index(x_test)
```

For simplicity, we use `SimpleLM` as the learner, which is a simple wrapper for `lm`.
 * `lm` is not compatible with `narx` as it requires a formula. 
 `SimpleLM` allows using a linear model without resorting to crafting formulas, similar to what `svm` from package `e1071` does.

Model one, without `xreg`:
```{r xreg1}
model = narx(x_train, SimpleLM, p = 2)
pred1 = forecast(model, h=10)
```

Model two, with correlated (good) `xreg`:
```{r xreg2}
model2 = narx(x_train, SimpleLM, p = 2, xreg = xreg)
pred2 = forecast(model2, xreg=xreg[ind_test])
```

Model three, with an uncorrelated (bad) `xreg`:
```{r xreg3}
model3 = narx(x_train, SimpleLM, p = 2, xreg = yreg)
pred3 = forecast(model3, xreg=yreg[ind_test])
```

The results:
```{r xregres}
rmse <- function(x,y) sqrt(mean((x-y)^2))

c(Err_without_xreg= rmse(pred1$mean, x_test),
  Err_with_xreg= rmse(pred2$mean, x_test),
  Err_with_bad_xreg= rmse(pred3$mean, x_test))
```

The smallest error is obtained from the model with correlated `xreg`.

### Notes

 * Some ML models, such as `SimpleLM`, require the same column names in training and testing to fit data.
 * Instead of `h` as the forecast horizon, `xreg` and its time-stamps are used.

